Metrics

### 2023-12-01 src/metrics.Rmd
Add cod plot and fix k for pccc

    # Calc k if pccc as it is diff for model combos
    if (identical(func, calc_pccc)) {
        out <- out %>%
            summarise(
                !!name := func(
                    actual,
                    prediction,
                    k = str_count(unique(Model), "_") + 1,
                    ...
                )
            )
    } else { # otherwise apply func
        out <- out %>%
            summarise(
                !!name := func(
                    actual,
                    prediction,
                    ...
                )
            )
    }

@@

# Bulk calc per model based on age and physician coding stage
calc_by_age_stage <- function(df, func, name = "Metric") {
    out <- list()

    # Calc overall metric
    out[[name]] <- calc_per_model(
        actual = df %>% pull(physician_cghr10),
        pred = df %>% select(ends_with("_cghr10"), -physician_cghr10),
        func = func
    ) %>% rename(
        {{ name }} := Metric
    )

    # Calc metric for va coding stage
    for (stage in c("is_agreed", "is_recon", "is_adj")) {

@@

        # Calc metric for stage
        out[[metric_name]] <- calc_per_model(
            actual = df_filter %>%
                pull(physician_cghr10),
            pred = df_filter %>%
                select(ends_with("_cghr10"), -physician_cghr10),
            func = func
        ) %>% rename(
             {{ metric_name }} := Metric
        )
    }

    # Calc pcc for each age group
    for (age_group in c("adult", "child", "neo")) {

@@

        # Calc metric for age
        out[[metric_name]] <- calc_per_model(
            actual = df_filter %>%
                pull(physician_cghr10),
            pred = df_filter %>%
                select(ends_with("_cghr10"), -physician_cghr10),
            func = func
        ) %>% rename(
             {{ metric_name }} := Metric
        )

@@

            # Calc metric for stage within age group
            out[[metric_name]] <- calc_per_model(
                actual = df_filter %>%
                    pull(physician_cghr10),
                pred = df_filter %>%
                    select(ends_with("_cghr10"), -physician_cghr10),
                func = func
            ) %>% rename(
                 {{ metric_name }} := Metric
            )
        }
    }

@@

```{r}
metrics$pccc <- calc_by_age_stage(df, calc_pccc, "PCCC")
metrics$pccc
```

@@

```{r}
metrics$csmf_acc <- calc_by_age_stage(df, calc_csmf_acc, "CSMF Accuracy")
metrics$csmf_acc
```

@@

# Plot boxplot
allplot <- alldata %>%
    ggplot(aes(x = Stage, y = PCCC)) +
    geom_boxplot(
        linewidth = 0.5,
        width = 0.25
    ) +
    geom_dotplot(
        binaxis = "y",
        binwidth = 0.03,
        dotsize = 0.3,
        fill = "white",
        color = "darkgray",
        stackdir = "center",
        stackratio = 2,
        position = position_nudge(x = -0.3)
    ) +
    geom_text(
        aes(label = `PCCC Min`),
        color = "#4d4d4d",
        size = 2,
        hjust = 1,
        position = position_nudge(y = -0.015, x = -0.3)
    ) +
    geom_text(
        aes(label = `PCCC Max`),
        color = "#4d4d4d",
        size = 2,
        position = position_nudge(y = 0.065, x = -0.3)
    ) +
    geom_text(
        aes(label = `PCCC Mid`),
        color = "#4d4d4d",
        size = 2,
        position = position_nudge(y = -0.065, x = -0.3)
    ) +
    stat_summary( # min pccc txt on boxplot
        geom = "text",
        fun = min,
        aes(label = sprintf("%1.2f", after_stat(y))),
        position = position_nudge(y = -0.025),
        size = 3
    ) +
    stat_summary( # max pccc txt on boxplot
        geom = "text",
        fun = max,
        aes(label = sprintf("%1.2f", after_stat(y))),
        position = position_nudge(y = 0.025),
        size = 3
    ) +
    geom_segment(aes(x = -Inf, y = 0, xend = Inf, yend = 0), arrow = arrow(length = unit(1, "cm"))) +
    labs(
        x = element_blank(),
        y = "PCCC (0=Low, 1=High)"
    ) +
    ylim(0.2, 0.8) +
    coord_flip() +
    theme_minimal() +
    theme(
        panel.grid = element_blank(),
        axis.line = element_line(color = "black"),
        axis.title.x = element_text(margin = margin(t = 5)),
        axis.title.y = element_text(margin = margin(r = 5)),
        axis.line.x = element_line(
            arrow = grid::arrow(length = unit(0.2, "cm"), ends = "both")
        ),
        axis.line.y = element_blank()
    )

@@

```{r}
```

@@

# Plot boxplot
ageplot <- agedata %>%
    ggplot(aes(x = `Age Group`, y = PCCC)) +
    geom_boxplot(
        linewidth = 0.5,
        width = 0.25
    ) +
    geom_dotplot(
        binaxis = "y",
        binwidth = 0.03,
        dotsize = 0.3,
        fill = "white",
        color = "darkgray",
        stackdir = "center",
        stackratio = 2,
        position = position_nudge(x = -0.3)
    ) +
    geom_text(
        aes(label = `PCCC Min`),
        color = "#4d4d4d",
        size = 2,
        hjust = 1,
        position = position_nudge(y = -0.015, x = -0.3)
    ) +
    geom_text(
        aes(label = `PCCC Max`),
        color = "#4d4d4d",
        size = 2,
        position = position_nudge(y = 0.055, x = -0.3)
    ) +
    geom_text(
        aes(label = `PCCC Mid`),
        color = "#4d4d4d",
        size = 2,
        hjust = 1,
        position = position_nudge(y = -0.015, x = -0.3)
    ) +
    stat_summary( # min pccc txt on boxplot
        geom = "text",
        fun = min,
        aes(label = sprintf("%1.2f", after_stat(y))),
        position = position_nudge(y = -0.025),
        size = 3
    ) +
    stat_summary( # max pccc txt on boxplot
        geom = "text",
        fun = max,
        aes(label = sprintf("%1.2f", after_stat(y))),
        position = position_nudge(y = 0.025),
        size = 3
    ) +
    geom_segment(aes(x = -Inf, y = 0, xend = Inf, yend = 0), arrow = arrow(length = unit(1, "cm"))) +
    labs(
        x = paste0("Agreed Records (n=", nagree, ", 100%)"),
        y = "PCCC (0=Low, 1=High)"
    ) +
    ylim(0.2, 0.8) +
    coord_flip() +
    theme_minimal() +
    theme(
        panel.grid = element_blank(),
        axis.line = element_line(color = "black"),
        axis.title.x = element_text(margin = margin(t = 5)),
        axis.title.y = element_text(margin = margin(r = 5)),
        axis.line.x = element_line(
            arrow = grid::arrow(length = unit(0.2, "cm"), ends = "both")
        ),
        axis.line.y = element_blank()
    )

>>>>

    # Calc k if pccc as it is diff for model combos
    if (identical(func, calc_pccc)) {
        out <- out %>%
            summarise(
                !!name := func(
                    actual,
                    prediction,
                    k = str_count(unique(Model), "_"),
                    ...
                )
            )
    } else { # otherwise apply func
        out <- out %>%
            summarise(
                !!name := func(
                    actual,
                    prediction,
                    ...
                )
            )
    }

@@

# Bulk calc per model based on age, physician coding stage, and cause of death
calc_by <- function(
        df,
        func,
        name = "Metric",
        by_age = TRUE,
        by_stage = TRUE,
        by_cod = TRUE,
        ...
    ) {
    out <- list()

    # Calc overall metric
    out[[name]] <- calc_per_model(
        actual = df %>% pull(physician_cghr10),
        pred = df %>% select(ends_with("_cghr10"), -physician_cghr10),
        func = func,
        ...
    ) %>% rename(
        {{ name }} := Metric
    )

    # Calc metric for va coding stage
    if (by_stage) {
        for (stage in c("is_agreed", "is_recon", "is_adj")) {

@@

            # Calc metric for stage
            out[[metric_name]] <- calc_per_model(
                actual = df_filter %>%
                    pull(physician_cghr10),
                pred = df_filter %>%
                    select(ends_with("_cghr10"), -physician_cghr10),
                func = func,
                ...
            ) %>% rename(
                 {{ metric_name }} := Metric
            )
        }
    }

    # Calc pcc for each age group
    if (by_age) {
        for (age_group in c("adult", "child", "neo")) {

@@

            # Calc metric for age
            out[[metric_name]] <- calc_per_model(
                actual = df_filter %>%
                    pull(physician_cghr10),
                pred = df_filter %>%
                    select(ends_with("_cghr10"), -physician_cghr10),
                func = func,
                ...
            ) %>% rename(
                 {{ metric_name }} := Metric
            )

@@

                # Calc metric for stage within age group
                out[[metric_name]] <- calc_per_model(
                    actual = df_filter %>%
                        pull(physician_cghr10),
                    pred = df_filter %>%
                        select(ends_with("_cghr10"), -physician_cghr10),
                    func = func,
                    ...
                ) %>% rename(
                     {{ metric_name }} := Metric
                )
            }
        }
    }

    # Calc metric by cause of death
    if (by_cod) {

        # Get unique causes of death
        causes <- df %>%
            select(ends_with("_cghr10")) %>%
            pivot_longer(
                everything(),
                names_to = "column",
                values_to = "cod"
            ) %>%
            distinct(cod) %>%
            filter(!is.na(cod)) %>%
            pull(cod)

        # Calc metric by cod
        for (cod in causes) {

            # Filter data for cod
            df_filter <- df %>% filter(physician_cghr10 == cod)
            metric_name = paste0(name, " COD ", cod)

            # Calc metric for cod
            out[[metric_name]] <- calc_per_model(
                actual = df_filter %>%
                    pull(physician_cghr10),
                pred = df_filter %>%
                    select(ends_with("_cghr10"), -physician_cghr10),
                func = func,
                ...
            ) %>% rename(
                 {{ metric_name }} := Metric
            )
        }
    }

@@

# Get num of unique causes
ncauses <- df %>%
    select(ends_with("_cghr10")) %>%
    pivot_longer(
        everything(),
        names_to = "column",
        values_to = "cod"
    ) %>%
    distinct(cod) %>%
    filter(!is.na(cod)) %>%
    pull(cod) %>%
    length

# Calc pccc for age, stage, and cod
metrics$pccc <- calc_by(df, calc_pccc, "PCCC", N = ncauses)
metrics$pccc
```

@@

```{r}
metrics$csmf_acc <- calc_by(
    df,
    calc_csmf_acc,
    "CSMF Accuracy",
    by_cod = FALSE
)
metrics$csmf_acc
```

@@

# Plot boxplot
allplot <- alldata %>%
    ggplot(aes(x = Stage, y = PCCC)) +
    geom_boxplot(
        linewidth = 0.5,
        width = 0.25
    ) +
    geom_dotplot(
        binaxis = "y",
        binwidth = 0.03,
        dotsize = 0.3,
        fill = "white",
        color = "darkgray",
        stackdir = "center",
        stackratio = 2,
        position = position_nudge(x = -0.3)
    ) +
    geom_text(
        aes(label = `PCCC Min`),
        color = "#4d4d4d",
        size = 2,
        hjust = 1,
        position = position_nudge(y = -0.015, x = -0.3)
    ) +
    geom_text(
        aes(label = `PCCC Max`),
        color = "#4d4d4d",
        size = 2,
        position = position_nudge(y = 0.065, x = -0.3)
    ) +
    geom_text(
        aes(label = `PCCC Mid`),
        color = "#4d4d4d",
        size = 2,
        hjust = 1,
        position = position_nudge(y = -0.015, x = -0.3)
    ) +
    stat_summary( # min pccc txt on boxplot
        geom = "text",
        fun = min,
        aes(label = sprintf("%1.2f", after_stat(y))),
        position = position_nudge(y = -0.025),
        size = 3
    ) +
    stat_summary( # max pccc txt on boxplot
        geom = "text",
        fun = max,
        aes(label = sprintf("%1.2f", after_stat(y))),
        position = position_nudge(y = 0.025),
        size = 3
    ) +
    labs(
        x = element_blank(),
        y = "PCCC (0=Low, 1=High)"
    ) +
    ylim(0.2, 0.8) +
    coord_flip() +
    theme_minimal() +
    theme(
        panel.grid = element_blank(),
        axis.line = element_line(color = "black"),
        axis.title.x = element_text(margin = margin(t = 5)),
        axis.title.y = element_text(margin = margin(r = 5)),
        axis.line.x = element_line(
            arrow = grid::arrow(length = unit(0.2, "cm"), ends = "both")
        ),
        axis.line.y = element_blank()
    )

@@

```{r, fig.height = 16, fig.width = 8}

# Calc total agreed
nagree <- df %>% filter(is_agreed == TRUE) %>% nrow

# Get unique cause columns
causes_col <- out %>% select(starts_with("PCCC COD ")) %>% names

# Create cod with counts
cod_remap <- list()
for (cod_col in causes_col) {

    # Get cod without prefix
    cod <- gsub("PCCC COD ", "", cod_col)

    # Calculate num and perc cases for cod
    ncod <- df %>% filter(physician_cghr10 == cod & is_agreed == TRUE) %>% nrow
    pcod <- (ncod / nagree) * 100
    pcod_label <- if (round(pcod) >= 1) round(pcod) else "<1"

    # Break cod into newlines if more than 3 words
    if (str_count(cod, "\\s+") > 3) {
        cod_label <- str_replace(cod, "(\\S+\\s+\\S+\\s+\\S+) ", "\\1\n")
    } else {
        cod_label <- cod
    }

    # Add label for cod
    cod_remap[[cod]] <- paste0(cod_label, "\n(n=", ncod, ", ", pcod_label, "%)")
}

# Prepare boxplot data
coddata <- out %>%
    select(
        Model,
        starts_with("PCCC COD ")
    ) %>%
    rename_at(
         vars(-Model),
         ~gsub("PCCC COD ", "", .)
    ) %>%
    filter(!str_detect(Model, "&")) %>% # remove multi models
    pivot_longer( # transform to long format
        cols = -Model,
        names_to = "COD",
        values_to = "PCCC"
    ) %>%
    mutate( # rename cod with counts
        "COD" = recode(`COD`, !!!cod_remap)
    ) %>%
    group_by(`COD`) %>%
    mutate( # Make values less than 0 equal to 0
        PCCC = if_else(PCCC <= 0, 0, PCCC)
    ) %>%
    mutate( # Add min, mid, and max model names to label
        "PCCC Min" = if_else(PCCC <= min(PCCC) * 1.03, paste0(Model, " (", round(PCCC, 2), ")"), NA),
        "PCCC Max" = if_else(PCCC >= max(PCCC) * 0.97, paste0(Model, " (", round(PCCC, 2), ")"), NA),
        "PCCC Mid" = if_else(is.na(`PCCC Min`) & is.na(`PCCC Max`), paste0(Model, "\n(", round(PCCC, 2), ")"), NA),
        "PCCC Min" = if_else(`PCCC` == min(PCCC), paste(na.omit(`PCCC Min`), collapse = "\n"), NA), # Combine into one row if close PCCC
        "PCCC Max" = if_else(`PCCC` == max(PCCC), paste(na.omit(`PCCC Max`), collapse = "\n"), NA)
    )

# Create cod order based on max pccc
codorder <- coddata %>%
    group_by(COD) %>%
    summarise("PCCC Max Value" = max(PCCC, na.rm = TRUE)) %>%
    select(COD, `PCCC Max Value`) %>%
    arrange(desc(`PCCC Max Value`)) %>%
    pull(COD)
coddata$COD <- factor(coddata$COD, levels = rev(codorder))

# Plot boxplot
codplot <- coddata %>%
    ggplot(aes(x = COD, y = PCCC)) +
    geom_boxplot(
        linewidth = 0.5,
        width = 0.25
    ) +
    geom_text(
        aes(label = `PCCC Min`),
        color = "#4d4d4d",
        size = 2.5,
        hjust = 1,
        position = position_nudge(y = -0.12)
    ) +
    geom_text(
        aes(label = `PCCC Max`),
        color = "#4d4d4d",
        size = 2.5,
        hjust = 0,
        position = position_nudge(y = 0.12)
    ) +
    stat_summary( # min pccc txt on boxplot
        geom = "text",
        fun = min,
        aes(label = sprintf("%1.2f", after_stat(y))),
        position = position_nudge(y = -0.06),
        size = 3
    ) +
    stat_summary( # max pccc txt on boxplot
        geom = "text",
        fun = max,
        aes(label = sprintf("%1.2f", after_stat(y))),
        position = position_nudge(y = 0.06),
        size = 3
    ) +
    labs(
        x = paste0("Agreed Records (n=", nagree, ", 100%) for ", length(causes_col), " Causes of Death"),
        y = "PCCC (0=Low, 1=High)"
    ) +
    ylim(-0.3, 1.25) +
    coord_flip() +
    theme_minimal() +
    theme(
        panel.grid = element_blank(),
        axis.line = element_blank(),
        axis.line.y = element_blank(),
        axis.title.x = element_text(margin = margin(t = 5)),
        axis.title.y = element_text(margin = margin(r = 5)),
        axis.line.x = element_line(
            color = "black",
            arrow = grid::arrow(length = unit(0.2, "cm"), ends = "both")
        )
    )

# Save the plot
codplot
ggsave("../manuscript/fig-perf-cod.pdf", plot = codplot, dpi = 300)
```

@@

# Plot boxplot
ageplot <- agedata %>%
    ggplot(aes(x = `Age Group`, y = PCCC)) +
    geom_boxplot(
        linewidth = 0.5,
        width = 0.25
    ) +
    geom_dotplot(
        binaxis = "y",
        binwidth = 0.03,
        dotsize = 0.3,
        fill = "white",
        color = "darkgray",
        stackdir = "center",
        stackratio = 2,
        position = position_nudge(x = -0.3)
    ) +
    geom_text(
        aes(label = `PCCC Min`),
        color = "#4d4d4d",
        size = 2,
        hjust = 1,
        position = position_nudge(y = -0.015, x = -0.3)
    ) +
    geom_text(
        aes(label = `PCCC Max`),
        color = "#4d4d4d",
        size = 2,
        position = position_nudge(y = 0.055, x = -0.3)
    ) +
    geom_text(
        aes(label = `PCCC Mid`),
        color = "#4d4d4d",
        size = 2,
        hjust = 1,
        position = position_nudge(y = -0.015, x = -0.3)
    ) +
    stat_summary( # min pccc txt on boxplot
        geom = "text",
        fun = min,
        aes(label = sprintf("%1.2f", after_stat(y))),
        position = position_nudge(y = -0.025),
        size = 3
    ) +
    stat_summary( # max pccc txt on boxplot
        geom = "text",
        fun = max,
        aes(label = sprintf("%1.2f", after_stat(y))),
        position = position_nudge(y = 0.025),
        size = 3
    ) +
    labs(
        x = paste0("Agreed Records (n=", nagree, ", 100%)"),
        y = "PCCC (0=Low, 1=High)"
    ) +
    ylim(0.2, 0.8) +
    coord_flip() +
    theme_minimal() +
    theme(
        panel.grid = element_blank(),
        axis.line = element_line(color = "black"),
        axis.title.x = element_text(margin = margin(t = 5)),
        axis.title.y = element_text(margin = margin(r = 5)),
        axis.line.x = element_line(
            arrow = grid::arrow(length = unit(0.2, "cm"), ends = "both")
        ),
        axis.line.y = element_blank()
    )
